Automated detection \& characterization of long-period transits in the \kepler\ archive

### 2016-01-08 document/ms.tex
dmv writing

% Notation for this paper.
\newcommand{\period}{{\ensuremath{P}}}
\newcommand{\rp}{{\ensuremath{R_\mathrm{P}}}}
\newcommand{\rate}{{\ensuremath{\Gamma}}}
\newcommand{\score}{{\ensuremath{S}}}

@@

For the purposes of this \paper, we select the 40,000 brightest G and K dwarfs
from the \kepler\ catalog using the following parameters:

        % m = (4200 <= stlr.teff) & (stlr.teff <= 6100)
        % m &= stlr.radius <= 1.15
        % m &= stlr.dataspan > 365.25*2.
        % m &= stlr.dutycycle > 0.6
        % m &= stlr.rrmscdpp07p5 <= 1000.
        % m &= stlr.kepmag < 15.

\begin{itemize}
{\item $4200\unit{K} \le T_\mathfm{eff} \le 6100\unit{K}$}

\end{itemize}

@@

\section{Estimated yield}\sectlabel{est}

Before the launch of the \kepler\ Mission, \citet{Yee:2008} predicted the
yield of single transit events based on the planet occurrence rates estimated
based on small catalog of radial velocity discoveries \citep{Butler:2006} and
a fit to their occurrence rate \citep{Cumming:2008}.
Using these early results and the pre-launch specifications of the \kepler\
Mission, \citet{Yee:2008} predicted that \kepler\ would discover $\sim 6$
single transit events in the full dataset.
Now that the Mission is complete and we have better estimates of the
occurrence rate and distribution of planets \citep[for example][]{Dong:2013,
Petigura:2013, Foreman-Mackey:2014, Dressing:2015}, we can update the
predicted yield for single transit events in the \kepler\ light curves.
To make this estimate, we extrapolate the distribution of large planets on
relatively short orbits out to longer periods, taking detection efficiency
and the survey and targeting properties into account.

We will base our estimate on an assumed model $Q_k(\rp,\,\period)$ for the
absolute probability of detecting a planet with radius \rp\ and period
\period\ orbiting the star $k$, and the occurrence rate distribution
\begin{eqnarray}
\rate (\rp,\,\period) &=& \frac{\dd N}{\dd\ln\rp\dd\ln\period} \quad,
\end{eqnarray}
the expected number of planets per star, per logarithmic radius, per
logarithmic period.
Given these two quantities, the expected number of single transits in the
\kepler\ data is given by
\begin{eqnarray}
N &=& \sum_{k=1}^K N_k
\end{eqnarray}
where the sum is over the $K$ stars in the sample and $N_k$ is the expected
number of observable transits around star $k$
\begin{eqnarray}\eqlabel{nk}
N_k &=& \int Q_k(\rp,\,\period) \, \rate (\rp,\,\period)
    \dd\ln\rp \dd\ln\period \quad.
\end{eqnarray}
The integral in \eq{nk} is over the full range of target parameters.

For the purposes of this discussion, we assume an approximate simple detection
efficiency model with three contributions: the geometric transit probability,
the temporal transit probability, and signal-to-noise ratio threshold of the
search technique.
Assuming circular orbits, the geometric transit probability is given by
\citep{Winn:2010}
\begin{eqnarray}\eqlabel{q-geom}
Q_k^\mathrm{(geom)}(\rp,\,\period) &=& \frac{R_k}{a} \\
&=& \left( \frac{4\,\pi^2}{G\,M_k} \right)^{1/3} \, R_k \, \period^{-2/3}
\end{eqnarray}
where $R_k$ and $M_k$ are the radius and mass of the star $k$ respectively.
The eccentricity distribution of these long-period planets will affect this
transit probability \citep{Kipping:2014} but this simple prescription should
be sufficient for a rough estimate.
For long-period orbits, the temporal transit probability will be given by
\begin{eqnarray}\eqlabel{q-time}
Q_k^\mathrm{(time)}(\rp,\,\period) &=& \frac{T_k}{\period}
\end{eqnarray}
where $T_k$ is the total time that \kepler\ spent observing the star $k$.
Finally, the detection threshold depends on the detailed sensitivity of the
search procedure but we will approximate it as a simple step function in
signal-to-noise ratio of the transit.
This contribution will be given approximately by
\begin{eqnarray}
Q_k^\mathrm{(detect)}(\rp,\,\period) &=& \left\{\begin{array}{ll}
1 & \mathrm{if}\,\left(\rp/R_k\right)^2 > f\,\sigma_k \\
0 & \mathrm{otherwise}
\end{array}\right.
\end{eqnarray}
where $\sigma_k$ is an estimate of the noise in light curve of star $k$ and
$f$ is the detection threshold for the method.

Combining the detection efficiency components, the integral from \eq{nk}
becomes
\begin{eqnarray}\eqlabel{nk-2}
N_k &=& \left( \frac{4\,\pi^2}{G\,M_k} \right)^{1/3} \, R_k \, T_k
    \int_{\rp_\mathrm{min}} ^{\rp_\mathrm{max}} \frac{\dd\rp}{\rp}
    \int_{\period_\mathrm{min}} ^{\period_\mathrm{max}}
        \period^{-8/3}\,\rate(\rp,\,\period) \dd\period
\end{eqnarray}
where all the integration limits are set by the target parameter space.
Because of the detection probability threshold, $\rp_\mathrm{min}$ can be no
smaller than $R_k\,\sqrt{f\,\sigma_k}$.
\citet{Dong:2013} used the catalog of short period transiting planets found
by \kepler\ to constrain a model for the occurrence rate of large planets of
the form
\begin{eqnarray}
\rate(\rp,\,\period) &=& C\,\left(\frac{\period}{10\,\mathrm{d}}\right)^\beta
\end{eqnarray}
in a set of radius bins.
Using this model, the integral in \eq{nk-2} becomes
\begin{eqnarray}
N_k &=& \left( \frac{4\,\pi^2}{G\,M_k} \right)^{1/3} \, R_k \, T_k \,
    \sum_{j=1}^J \frac{C_j}{(10\,\mathrm{d})^{\beta_j}}\,
    \ln \left( \frac{\rp_{\mathrm{max},j}}{\rp_{\mathrm{min},j}} \right) \,
    \left[ \frac{{\period_\mathrm{min}}^{\beta_j - 5/3}}{5/3-\beta_j} \right]
\end{eqnarray}
where the sum is over the $J$ radial bins studied by \citet{Dong:2013}.
It's important to note that the model used by \citet{Dong:2013} is given in
base-10 logarithms so the units must be converted to natural logarithms as
appropriate.

Assuming the stellar parameters provided by the NASA Exoplanet
Archive\footnote{We downloaded the \texttt{q1\_q16\_stellar} table from
\url{http://exoplanetarchive.ipac.caltech.edu/} on 2015-04-03.}
\citep{Huber:2014} and approximating the stellar noise using the 15-hour CDPP
\citep{Christiansen:2012}, \fig{predict} shows the extrapolated number of
transiting planets with $1500\,\mathrm{d} < \period < 5000\,\mathrm{d}$ based
on the \citet{Dong:2013} power-law model.
For a detection threshold $f\sim10$, the expected number of single transit
events in the \kepler\ light curves is $\sim 60$.

This results is much more optimistic than the pre-launch estimate from
\citet{Yee:2008} for a few reasons.
One effect is that, the \kepler\ Mission ran for more than four
years---longer than the fiducial Mission goal---and $\sim190,000$ stars were
targeted for nearly the full baseline instead of the original 100,000.
\dfmtodo{Explain the other reasons.}

>>>>

% Notation for this paper.
\newcommand{\period}{{\ensuremath{P}}}
\newcommand{\unit}[1]{{\ensuremath{\,\mathrm{#1}}}}

@@

For the purposes of this \paper, we select the $\sim40,000$ brightest and
quietest G and K dwarfs from the \kepler\ catalog using the following
parameters:
\begin{itemize}
{\item $4200\unit{K} \le T_\mathrm{eff} \le 6100\unit{K}$,}
{\item $R_\star \le 1.15\,R_\odot$,}
{\item $K_p \le 15\unit{mag}$, and}
{\item $\mathrm{CDPP}_{7.5\unit{hrs}} \le 1000\unit{ppm}$.}
\end{itemize}
We also restrict the sample to only include targets with a data span of more
than two years and a duty cycle of better than $0.6$.
Since the \kepler\ data have been searched for shorter period planets
\dfmtodo{CITE Coughlin} and eclipsing binaries \dfmtodo{CITE EB catalog} than
targeted by this project, we censor these previously known targets.
For the known EBs, we simply remove these targets from the candidate list.
For the confirmed and candidate KOIs from \dfmtodo{Coughlin}, we remove data
within 2 transit durations of the known candidates.

@@

\section{Supervised classification}
